---
title: Analysis of the TOP Framework UX Survey Results
date: last-modified
authors:
  - name: Christoph Beger
    orcid: 0000-0002-1166-0368
    email: christoph.beger@medizin.uni-leipzig.de
    corresponding: true
    affiliations:
      - name: Leipzig University
        department: Institute of Medical Informatics, Statistics and Epidemiology
        city: Leipzig
        country: Germany
        role: author
format:
  typst:
    toc: true
    section-numbering: 1.1.1
    echo: false
    bibliography: references.bib
    bibliographystyle: elsevier-vancouver
---

```{=typst}
#show figure: set block(breakable: true)
#set page(header: align(right + horizon, [Analysis of the TOP Framework UX Survey Results]))
```

# Introduction

This script processes and analyses data collected from an online survey evaluating user experience (UX) with the TOP
Framework [@begerModeldrivenExecutionPhenotype2023; @begerModellingAdverseEvents2023].
The User Experience Questionnaire (UEQ, version 11) [@schreppMeasuringUserExperience2021] was used to assess
the user experience.
This includes steps for data preprocessing, calculation of summary statistics for UX-related scales, and generation of
tables and visualizations to facilitate interpretation of the survey findings.
The script is intended for research and reporting purposes, using survey data provided in CSV format.

The analysis supports understanding of user perceptions and experiences with the TOP Framework, providing insights for
further development and improvement.

Analysis were performed with `r R.version.string` [@rcoreteamLanguageEnvironmentStatistical2024].

```{r setup, include=FALSE}
library(tidyverse)
library(gtsummary)
library(gt)
library(rstatix)
library(purrr, include.only = "partial")
```

```{r data-preparation}
theme_gtsummary_mean_sd()
style_number_2digits <- partial(style_number, digits = 2)

data <- read.csv("data/NWGTOP_DATA_2025-03-28_1019.csv")

result <- data |>
  select(record_id, matches("q_\\d+")) |>
  filter(if_any(q_1:q_26, ~ !is.na(.x))) |>
  mutate(
    # scales are ordered randomly and must be standardised to the interval [-3; 3]
    across(c(q_1:q_2, q_6:q_8, q_11, q_13:q_16, q_20, q_22, q_26), ~ .x - 4),
    across(c(q_3:q_5, q_9:q_10, q_12, q_17:q_19, q_21, q_23:q_25), ~ 4 - .x),
    # calculate mean values for each scale
    attractiveness = mean(c(q_1, q_12, q_14, q_16, q_24, q_25), na.rm = TRUE),
    perspicuity = mean(c(q_2, q_4, q_13, q_21), na.rm = TRUE),
    efficiency = mean(c(q_9, q_20, q_22, q_23), na.rm = TRUE),
    dependability = mean(c(q_8, q_11, q_17, q_19), na.rm = TRUE),
    stimulation = mean(c(q_5, q_6, q_7, q_18), na.rm = TRUE),
    novelity = mean(c(q_3, q_10, q_15, q_26), na.rm = TRUE),
    attractiveness_quality = attractiveness,
    pragmatic_quality = mean(c(perspicuity, efficiency, dependability)),
    hedonic_quality = mean(c(stimulation, novelity)),
    .by = record_id
  ) |>
  select(-starts_with("q_"))
```

# Results

```{r}
#| label: tbl-summary
#| tbl-cap: |
#|   Summary statistics of the TOP Framework UX survey results.
#|   Data collected from an online survey evaluating user experience with the TOP Framework.

result |>
  tbl_summary(
    include = -record_id,
    type = everything() ~ "continuous2",
    missing = "no",
    statistic = all_continuous() ~ c(
      "{mean} ({sd})",
      "{min} - {max}",
      "{median} ({p25}, {p75})"
    ),
    label = list(
      attractiveness_quality ~ "attractiveness quality",
      pragmatic_quality ~ "pragmatic quality",
      hedonic_quality ~ "hedonic quality"
    )
  ) |>
  add_ci(style_fun = list(everything() ~ style_number_2digits)) |>
  as_gt() |>
  tab_row_group("grouped scales", rows = 25:36) |>
  tab_row_group("scales", rows = 1:24) |>
  cols_width(1 ~ cm(2))
```

```{r}
#| label: tbl-t-test
#| tbl-cap: t-test results for the TOP Framework UX survey scales.

result |>
  pivot_longer(attractiveness:novelity) |>
  group_by(name) |>
  t_test(value ~ 1, mu = 0.8, detailed = TRUE) |>
  add_significance() |>
  select(name, n, estimate, alternative, conf.low, conf.high, p, p.signif) |>
  gt() |>
  fmt_number(-n, decimals = 3) |>
  cols_width(1 ~ cm(2))
```

```{r}
#| label: fig-scale-means
#| fig-cap: |
#|   Mean values of the UX scales with confidence intervals.
#|   The scales are colour-coded based on their performance: red for poor, yellow for acceptable, and green for good.
#| fig-width: 10
#| fig-height: 5

result |>
  summarise(
    across(
      c(attractiveness, perspicuity, efficiency, dependability, stimulation, novelity),
      .fns = list(mean = mean, margin = ~ qt(0.975, df = n() - 1) * sd(.) / sqrt(n()))
    )
  ) |>
  tidyr::pivot_longer(
    everything(),
    names_sep = "_",
    names_to = c("scale", "property"),
  ) |>
  tidyr::pivot_wider(
    names_from = property,
    values_from = value
  ) |>
  mutate(scale = factor(
    scale,
    c("attractiveness", "perspicuity", "efficiency", "dependability", "stimulation", "novelity")
  )) |>
  ggplot(aes(scale, mean)) +
  annotate(
    geom = "rect",
    xmin = -Inf,
    xmax = Inf,
    ymin = c(-Inf, -0.8, 0.8),
    ymax = c(-0.8, 0.8, Inf),
    fill = c("#fcbbbbff", "#fcf4b1ff", "#bafcbaff")
  ) +
  geom_hline(yintercept = c(-0.8, 0.8), linetype = rep("dashed", 2)) +
  geom_hline(yintercept = -2:2, color = "gray60", linetype = "solid") +
  geom_bar(stat = "identity", width = 0.8, fill = "gray50") +
  geom_errorbar(
    aes(ymin = mean - margin, ymax = mean + margin),
    width = 0.3,
    linewidth = 1
  ) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 15),
    axis.title = element_text(size = 20, colour = "gray30"),
  ) +
  xlab(NULL) +
  coord_cartesian(ylim = c(-1, 2.5))

ggsave("data/scale-means.png", width = 10, height = 5, dpi = 600)
```